```{r, echo=FALSE}
htmltools::img(src = knitr::image_uri("Pictures_and_diagrams/WCS_LOGOTYPE.png"), 
               alt = 'logo', 
               style = 'position:fixed; 
                        top:0;
                        right:0; 
                        padding:70px;',
               width = "300px", 
               heigth = "300px")


htmltools::img(src = knitr::image_uri("Pictures_and_diagrams/wcpa-logo.png"), 
               alt = 'logo', 
               style = 'position:fixed; 
                        top:90;
                        right:0; 
                        padding:70px;',
               width = "300px", 
               heigth = "300px")

```

# Wildlife Health Data Model Dictionary {.unnumbered}


```{r table summarizing the different sources, eval=FALSE, echo=F, message=FALSE, warning=FALSE}

library(kableExtra)

data <- data.frame(
  Unit = 
    c("Source",
      "Source Record",
      "Source Record\nCarcass",
      "Source Record\nSpecimen",
      "Implications"),
  Observation_Source = 
    c("Animals y of the same group and species",
      "Animals y of the same group and species at time",
      "NA",
      "Tissues from animals y of the same group and species collected at time t",
      "Single source record per source at time t"),
  Animal_Source = 
    c("Individual Animal y",
      "Individual Animal y at time t",
      "Carcass of individual animal y collected at time t",
      "Tissue collected from individual animal y at time t",
      "Single source record per source at time t"),
  Environmental_Source = 
    c("Collection site  y with tissue of types x,w, and z",
      "Collection site y with tissue of types x,w, and z at time t",
      "NA",
      "Tissue of type x,w, or z collected from collection site y at time t",
      "Single source record per source at time t"),
  Arthropod_Source = 
    c("Collection site y",
      "Arthropods from collection site y at time t divided by species",
      "NA",
      "Arthropods collected from collection site y at time t divided by species and divided in sub-groups (“pools”)",
      "Multiple source records per source (1 by arthropod species)")
)

kable(data) |> kable_paper()


```


```{r load packages, echo=FALSE, eval=T, message=FALSE, cache=F, warning=FALSE}
#| tbl-colwidths: false



# ------------------#
# Create Dictionary #
# ------------------#

library(kableExtra)
library(tidyverse)
library(gt)
library(reactable)

# #tables in the .xml of the db diagram
# fields_by_table<-split(full_table, full_table$Table_Name)


# Example entries
# data_dictionary<-vector(mode = "list")
  
```

## Project

```{r project, echo=FALSE, eval=T, message=FALSE, cache=F}

source("dd_project.R")

project

```

## Surveillance Activity 

```{r surveillance objective, echo=FALSE, eval=T, message=FALSE, cache=F}

source("dd_surv_act.R")

surv_activity
```

### Surveillance Activity - Group Source 

```{r group source metadata, echo=FALSE, eval=T, message=FALSE, cache=F}

source("dd_surv_act_obs_source.R")

group_source

```

#### Surveillance Activity - Group Source Records Clustering

```{r group source clustering metadata, echo=FALSE, eval=T, message=FALSE, cache=F}

source("dd_surv_act_clusters_group_source_records.R")

cluster_group_source_records

```

### Surveillance Activity - Animal Source 

```{r animal source metadata, echo=FALSE, eval=T, message=FALSE, cache=F}

source("dd_surv_act_animal_source.R")

animal_source

```

#### Surveillance Activity - Animal Source Records Clustering

```{r animal source clustering metadata, echo=FALSE, eval=T, message=FALSE, cache=F}

source("dd_surv_act_clusters_animal_source_records.R")

cluster_animal_source_records

```

### Surveillance Activity - Environmental Source

```{r environmental source metadata, echo=FALSE, eval=T, message=FALSE, cache=F}

source("dd_surv_act_env_source.R")

environmental_source

```

#### Surveillance Activity - Environmental Source Records Clustering

```{r environmental source clustering metadata, echo=FALSE, eval=T, message=FALSE, cache=F}

source("dd_surv_act_clusters_environmental_source.R")

cluster_environmental_source_records

```

### Surveillance Activity - Arthropod Source

```{r arthropod source metadata, echo=FALSE, eval=T, message=FALSE, cache=F}

source("dd_surv_act_arthro_source.R")

arthro_source

```

#### Surveillance Activity - Arthropod Source Records Clustering

```{r arthropod source clustering metadata, echo=FALSE, eval=T, message=FALSE, cache=F}

source("dd_surv_act_clusters_arthropod_source_records.R")

cluster_arthropod_source_records

```

### Surveillance Activity - Location Clustering

```{r location clustering metadata, echo=FALSE, eval=T, message=FALSE, cache=F}

source("dd_surv_act_clusters_location.R")

cluster_locations

```

### Surveillance Activity - Event Clustering

```{r event clustering metadata, echo=FALSE, eval=T, message=FALSE, cache=F}

source("dd_surv_act_clusters_event.R")

cluster_events

```



## Outbreak

```{r outbreak, echo=FALSE, eval=T, message=FALSE, cache=F}

source("dd_outbreak.R")

outbreak

```

## Field Visit

```{r field visit, echo=FALSE, eval=T, message=FALSE, cache=F}

source("dd_field_visit.R")

field_visit

```

## Location

```{r location, echo=FALSE, eval=T, message=FALSE, cache=F}

source("dd_location.R")

location

```

### Location Clustering

```{r location clustering, echo=FALSE, eval=T, message=FALSE, cache=F}

source("dd_location_clustering.R")

location_clustering

```

## Event

```{r event, echo=FALSE, eval=T, message=FALSE, cache=F}

source("dd_event.R")

event

```

### Event Clustering

```{r event clustering, echo=FALSE, eval=T, message=FALSE, cache=F}

source("dd_event_clustering.R")

event_clustering

```

## Collection

```{r collection, echo=FALSE, eval=T, message=FALSE, cache=F}

source("dd_collection.R")

collection

```

## Source 

```{r source, echo=FALSE, eval=T, message=FALSE, cache=F}

source("dd_source.R")

source

```

### Group Source 

```{r group source, echo=FALSE, eval=T, message=FALSE, cache=F}

source("dd_group_source.R")

group_source

```

### Animal Source 

```{r animal source, echo=FALSE, eval=T, message=FALSE, cache=F}

source("dd_animal_source.R")

animal_source

```

## Source Record

### Group Source Record

```{r group source record, echo=FALSE, eval=T, message=FALSE, cache=F}

source("dd_group_source_records.R")

group_source_record

```

#### Group Source Record - Clustering

```{r group source record clustering, echo=FALSE, eval=T, message=FALSE, cache=F}

source("dd_group_source_record_clustering.R")

group_source_record_clustering

```

### Animal Source Record

```{r animal source record, echo=FALSE, eval=T, message=FALSE, cache=F}

source("dd_animal_source_record.R")

animal_source_record

```

#### Animal Source Record - Vaccination

```{r vaccination, echo=FALSE, eval=T, message=FALSE, cache=F}

source("dd_vaccination.R")

animal_source_record_vaccination

```

#### Animal Source Record - Carcass

```{r carcass, echo=FALSE, eval=T, message=FALSE, cache=F}

source("dd_animal_source_record_carcass.R")

animal_source_record_carcass

```

#### Animal Source Record - Necropsy

```{r necropsy, echo=FALSE, eval=T, message=FALSE, cache=F}

source("dd_animal_source_record_necropsy.R")

animal_source_record_necropsy

```

#### Animal Source Record - Clustering

```{r animal source record clustering, echo=FALSE, eval=T, message=FALSE, cache=F}

source("dd_animal_source_record_clustering.R")

animal_source_record_clustering

```

### Environmental Source Record

```{r environmental source record, echo=FALSE, eval=T, message=FALSE, cache=F}

source("dd_environmental_source_record.R")

environmental_source_record

```

#### Environmental Source Record - Clustering

```{r environmental source record clustering, echo=FALSE, eval=T, message=FALSE, cache=F}

source("dd_environmental_source_record_clustering.R")

environmental_source_record_clustering

```

### Arthropod Source Record

```{r arthropod source record, echo=FALSE, eval=T, message=FALSE, cache=F}

source("dd_arthropod_source_record.R")

arthropod_source_record

```

#### Arthropod Source Record - Clustering

```{r Aathropod source record clustering, echo=FALSE, eval=T, message=FALSE, cache=F}

source("dd_arthropod_source_record_clustering.R")

arthropod_source_record_clustering

```

## Specimen

```{r specimen, echo=FALSE, eval=T, message=FALSE, cache=F}

source("dd_specimen.R")

specimen

```



## Storage

```{r storage, echo=FALSE, eval=T, message=FALSE, cache=F}

source("dd_storage.R")

storage

```

## Export

```{r export, echo=FALSE, eval=T, message=FALSE, cache=F}

source("dd_export.R")

export

```

## Diagnostic

```{r diagnostic, echo=FALSE, eval=T, message=FALSE, cache=F}

source("dd_diagnostic.R")

diagnostic

```

### Diagnostic Product

```{r diagnostic product, echo=FALSE, eval=T, message=FALSE, cache=F}

source("dd_diagnostic_product.R")

diagnostic_product

```

## Laboratory

```{r laboratory, echo=FALSE, eval=T, message=FALSE, cache=F}

source("dd_laboratory.R")

laboratory

```

## Interpretation 

### Diagnostic Interpretation

```{r diagnostic interpretation, echo=FALSE, eval=T, message=FALSE, cache=F}

source("dd_diagnostic_interpretation.R")

diagnostic_interpretation

```

### Specimen Interpretation

```{r specimen interpretation, echo=FALSE, eval=T, message=FALSE, cache=F}

source("dd_specimen_interpretation.R")

specimen_interpretation

```

### Source Record Interpretation

```{r source record interpretation, echo=FALSE, eval=T, message=FALSE, cache=F}

source("dd_source_record_interpretation.R")

source_record_interpretation

```

