<!-- # Source and Source Records {.unnumbered} -->

### General Overview

In the data model, a **Source** is a unit of interest that can be observed, provide **Specimens** for analysis or that can be observed-only. The data model manages four types of **Sources**: i) a co-specific group of animals (*Group*), ii) individual animals (*Animal*), iii) sites where arthropods can be obtained from (*Arthropod*), and iv) sites where biotic and abiotic tissue not linked to any other type of Source can be obtained from (*Environmental*). A **Group Source** could correspond to bats roosting at cave  X sourcing guano, an **Animal Source** could be a collared animal, an **Arthropod Source** could be a site where mosquito larvae are collected from, and an **Environmental Source** could be a pond where water, air, and sediment are collected from or feces found in the field whose animal of origin is unknown.  **Sources** contain time-independent data only, such as the species of an animal in the case of a *Group* or *Animal* **Sources** (more below).

All **Sources** can potentially be identified and tracked over time if needed. However, there are limitations, *Environmental* and *Arthropod* **Sources** are site-specific, making them easier to track across different time points. *Animal* and *Group* **Sources** may not always be individually identifiable, which can prevent tracking them over time.

**A Source collected, or captured at time *t* can be linked to Events through a Point Collection or not (see previous section) via a Source Record, which represents the Source sampled or observed at a specific time *t**.  **Source Records** contain only time-dependent data, such as health status at time *t* for *Group* and *Animal* Sources** or the number of arthropods trapped at time *t* for *Arthropod Sources*.

**Individually identified Sources can tracked across multiple Events, even across different Surveillance Activities**. For example, a bat (**Source**) captured in a mist net (**Collection**) deployed at a roost at time *t* (**Event**), individually identified, and later recaptured in another **Event** *t′*. This bat is linked to both the original capture **Event** and the recapture **Event**, with a unique **Source Record** for each occurrence.

**Sources that are not identified will be linked as a Source Record to the Event where they were observed or captured only**. For example, a bat (Source) that is captured (Collection Point) and sampled but not marked or a dead animal not identified found by a ranger during a patrol. The bat and the animal found are linked to a single Event, the one they were capture or observed. 

**Sources found opportunistically are still linked to an Event, even if no effort is involved**. For example, a member of the public reports a dead animal (**Source Record**) found at time *t* (**Source Record**) at a beach (**Event**).

**Sources Records** can provide between zero (e.g., observation only) up to indefinite number of **Specimens** to conduct **Diagnostic** tests. 

### Group Source

**A Group Source is a unit of co-specific individuals (animals of the same species) that belong to a common entity, such as a herd, roost, area, site, farm, cage, stall, or enclosure, forming a single epidemiological unit**.  **Group Sources** can be observed, captured, diagnosed as a full unit, and provide  **Specimens ** collectively at a given time *t* when the **Event** occurs. 

Key attributes of a **Group Source** include:

* Species
* Group Source ID
* Group Source Cross Reference ID
* Collection Cross Reference Origin

The primary purpose of a **Group Source** is to record individuals at the species level rather than tracking each animal individually. This approach is useful when herds are the unit of interest, when protected area rangers document multiple animals of the same species during a health event, or when animals in a market are grouped in cages or stalls without individual identification. For example, if a single cage contains animals from two different species and they are not tracked individually, they can be recorded as two separate **Group Sources**—one per species.

#### Group Source Record

**A Group Source Record documents the animals within a Group Source at time *t*, categorized by sex, age, and health status (e.g., healthy, injured, sick, or dead)**. Additional attributes include:

* Observed anomalies
* Potential cause of disease
* Potential cause of death

Since properties of a **Group Source Record** are recorded at the group level, multiple options can be reported for these attributes. For example, if a **Group Source Record** includes three dead animals, several potential causes of death may be listed. However, it is not possible to determine how these potential causes or other properties are distributed across individuals within the **Group Source Record**—only that they were present in at least one individual.

A **Group Source Record can include a mix of dead, diseased, injured, and healthy individuals of the same species, or it may consist solely of one category if all individuals share the same status (e.g., all are dead)**. Additionally, even a **Group Source Record** consisting only of healthy animals - only can be part of a Health Event, depending on the **Event**’s definition.

A **Group Source Record may contain a single individual**. For example, rangers patrolling a protected area might find one dead animal of species X and two dead animals of species Y at the same site (**Event**). The animals can be documented as distinct **Group Source Records** (one group per species).

An **Event** can include multiple ***Group Source Records** of the same species. For example, if a vendor at a market keeps animals of the same species in two separate cages, and the **Event** is a vendor, then each cage could be considered a distinct unit. Consequently, there would be two **Group Source Records** for the same species under the same vendor (i.e., within the same **Event**).

A **Group Source** can be directly used for a **Diagnostic** (e.g., assessing the body condition of a specific herd using a standardized metric). The data model supports **Diagnostics** applied to an entire group rather than just a **Specimen** taken from that **Group Source**. **Specimens** from **Group Sources** can be stored, exported, and used for **Diagnostics**. On the contrary,**Group Source** cannot have a collective necropsy. This is because their collection and handling present an opportunity to gather data at the individual level and recorded as an **Animal Source**. Carcasses of animals of a **Group Source** collected or taken for Necropsy must be converted into **Animal Sources**. These **Animal Sources** are recorded as former members of the **Group Source** (see the Complexities section for further details).

For similar reasons, **Group Sources** cannot be removed from the field. If dead animals from a **Group Source** are collected, each must be documented as an **Animal Sources** originating from the corresponding **Group Source** whose **Carcass** is taken (see **Animal Sources** and **Carcass** below). 

The database does not accept live animals from **Group Sources** be taken ex-situ.

An animal of species X recorded in a **Group Source Record** must not be duplicated as an **Animal Source Record**, and vice versa. If an **Event** contains both a **Group Source Record** and an **Animal Source Record** of species X, the total number of animals of that species in the **Event** is the sum of those in the **Group Source Record** and the individual(s) recorded separately as **Animal Source Record** (e.g., the individuals in a group of animals of species X plus a carcass collected of the same species that would have belonged to the count of dead animals of the **Group Source Record** if not collected).

For example, consider a herd of 20 cows illegally raised in a protected area whose health is assessed at time *t*. If two cows are sampled, the two sampled cows should be documented as two separate **Animal Source Records** (one per cow), while the remaining 18 cows are recorded as a single **Group Source Record**, categorized by sex, age, and health status. In this scenario, the total number of cows is 20—the 18 recorded in the **Group Source Record** plus the two documented as **Animal Source Records**. The "herd" identity of these 20 cows split in three **Source Records** can be maintained using a cluster (see Complexities). If, however, the **Group Source Record** incorrectly includes all 20 cows while two of them are also recorded as **Animal Source Record**, the total count would incorrectly sum to 22.

### Animal Source

**An Animal Source represents a single individual**. **Animal Sources** can be observed, captured, tested, necropsied, and provide **Specimens **at time *t*, directly or during a **Necropsy**.

Key attributes of a **Animal Source** include:

* Species
* Animal Source ID
* Animal Source Cross Reference ID
* Sex

For example, in a live animal market, if data is collected at the individual level, each animal—whether of the same or different species—must be recorded as a separate **Animal Source**. In contrast, if animals are documented collectively (e.g., by species in a shared cage), they would be classified under a **Group Source** rather than as individual **Animal Sources**.

Any past marking codes assigned of the animal are considered immutable and are recorded as part of the **Animal Source** data.

#### Animal Source Record

An **Animal Source Record** represents an individual **Animal Source** at time *t*. 

Key attributes of a **Animal Source Record** include:

* Age
* Observed Health Status
* Observed Anomalies
* Potential Causes of Disease or Death

Some properties of an **Animal Source Record** allow single or multiple selections. For instance, an animal's health status is recorded as one category (e.g., "live healthy" or "live sick"), whereas multiple anomalies may be reported for the same individual at time *t* (e.g., wounds, hair loss, diarrhea).

An **Animal Source Record** can be categorized as dead, diseased, injured, or healthy at time *t*. A live, healthy **Animal Source Record can still be part of an **Event**, depending on the Event's definition** (e.g., an apparent healthy animal of species X observed near dead animals of species X, Y, and Z, or captured for **Specimen** collection).

An **Animal Source Record** can be used directly for a **Diagnostic** (e.g., performing X-rays on a live animal at time *t*). The data model supports **Diagnostics** applied to the individual itself, rather than only to **Specimens** collected from the animal.

Live **Animal Sources** cannot be removed from the field (ex-situ). Only **Carcasses** of **Animal Sources** can be transported, stored, an necropsied in a facility.

An individual of species X recorded as part of a **Group Source Record** cannot also be included as an **Animal Source Record**, and vice versa. If an **Event** contains both a **Group Source Record** of species X and an **Animal Source Record** of species X, then the total number of animals of species X at the **Event** is the sum of the individuals recorded in the **Group Source Record** and the single individual recorded as an **Animal Source Record**.

For example, consider a herd of 20 cows illegally raised in a protected area, where their health is assessed at time *t*. If two cows are sampled individually, they are recorded as two separate **Animal Source Record** (one per sampled cow). The remaining 18 cows are documented as one **Group Source Record**, where the total number of individuals is categorized by sex, age, and health status. Thus, the total number of cows at the **Event** is 20 (18 from the **Group Source Record** + 2 from the two **Group Source Records**). To maintain the herd identity, a cluster can be used (see Complexities). If the **Group Source Record** initially included all 20 cows and two additional **Animal Source Record** were created for the sampled individuals, the total number of cows would incorrectly appear as 22—which must be avoided.

#### Vaccination

The data model allows the inclusion of vaccinations administered to an **Animal Source** at the time of capture or immobilization *t*. Multiple vaccination records can be added to an **Animal Source Record** as needed, ensuring that vaccination history accumulates for the corresponding **Animal Source**.

#### Carcass

Each **Carcass** recorded in the database originates from a single **Animal Source** and is collected at a specific time *t*. Therefore, a **Carcass** is always linked to one, and only one, **Animal Source Record**.

Key attributes of a **Carcass** include:

* Decomposition Condition
* Storage Ex-situ
* Storage during Transport
* Owner
* Availability

Just like **Group Source** carcasses, the carcass of an **Animal Source** may undergo through multiple shipments among different facilities and changes in storage over time. Both, the initial storage and changes in storage as well as shipments can be added cumulatively to the Animal Source, so the full history is available.

Carcasses do not provide **Specimens** directly; rather, **Specimens** are collected through the associated dead **Animal Source Record**. The origin of **Specimens** (primary necropsy, secondary necropsy, field, etc.) is tracked. Similarly, a **Carcass** is not directly used for **Diagnostics**. any diagnostic procedures that involve a full carcass, such as X-rays, applied to the corresponding dead **Animal Source Record** and documented accordingly.

#### Necropsy

A **Necropsy** is performed on a specific **Animal Source Carcass**. A **Necropsy** can take place is either at the time the dead animal is found (field necropsy) or after the **Carcass** has been collected in a facility.

Key attributes of a **Carcass** include:

* Necropsy Identifier
* Necropsy Cross Reference Identifier
* Necropsy Date
* Findings per system
* Availability

Necropsies can be classified as partial field necropsy, full necropsy, primary, and secondary. A primary necropsy typically begins with an intact **Carcass** that has not been previously examined in a facility. In contrast, a secondary necropsy is usually performed by a veterinary pathologist using either photographs from the primary **Necropsy** or conducting the **Necropsy** again.

A **Necropsy** does not provide **Specimens** directly; rather, **Specimens** are provided through the dead **Animal Source Carcass**.

The **Necropsy** findings are reported through text in fields such as "Gastrointestinal System", "Reproductive System", etc. The list of anatomical components encompassed by each of these elements or systems is provided in the table below and they follow the National Agricultural Library Thesaurus Concept Space (NALT), a controlled vocabulary of terms related to agricultural, biological, physical and social sciences (https://lod.nal.usda.gov/nalt/en/). 

##### Mouth and Beak

```{r necropsy mouth and beak, echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE, cache=F}

library(gt)
library(dplyr)

# source("Necropsy_anatomy/Necropsy_topology_mouth_beak.R")

# saveRDS(mouth_beak_table, "Necropsy_anatomy/Tables for manual/mouth.RDS")
mouth_beak_table <- readRDS("Necropsy_anatomy/Tables for manual/mouth.RDS")

# mouth_beak_table <-
# mouth_beak_table |> 
#   mutate(Topography = stringr::str_to_sentence(Topography)) |> 
#   mutate(Alternative_name = stringr::str_to_sentence(Alternative_name)) |> 
#   mutate(Definition_NALT = ifelse(Definition_NALT == "", NA, Definition_NALT)) |> 
#   arrange(Topography) |> 
#   kable("html") |>
#   kable_styling(
#     bootstrap_options = c("striped", "hover"),
#     full_width = FALSE,
#     font_size = 8
#   ) |>
#   column_spec(1, width = "22%") |>  # Topography
#   column_spec(2, width = "28%") |>  # Alternative_name
#   column_spec(3, width = "28%") |>  # MESH_URI
#   column_spec(4, width = "30%") |>  # Definition_NALT
#   # column_spec(5, width = "30%") |>  # Definition_MESH
#   # column_spec(6, width = "22%") |>  # Alternative_name
#   row_spec(0, bold = TRUE, font_size = 12) |>  # Header row styling
#   scroll_box(width = "740px", height = "1000px")
  


  mouth_beak_table %>% bind_rows() |> 
  gt::gt() %>%
  gt::tab_options(table.font.size = 12) %>%
  gt::cols_width(Topography ~ gt::pct(22),
                 Alternative_name ~ gt::pct(20),
                 Definition_NALT ~ gt::pct(30)) %>%
  gt::tab_style(
    style = cell_text(size = px(15), weight = "bold"),
    locations = cells_column_labels())


```

##### Eyes

```{r necropsy eyes, echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE, cache=F}

# source("Necropsy_anatomy/Necropsy_topology_eye.R")

# saveRDS(eye_table, "Necropsy_anatomy/Tables for manual/eye.RDS")
eye_table <- readRDS("Necropsy_anatomy/Tables for manual/eye.RDS")

# rbind(
#   eye_table,
#   c("Eyelashes", NA, NA, NA)) |> 
#   arrange(Topography) |> 
#   kable("html") |>
#   kable_styling(
#     bootstrap_options = c("striped", "hover"),
#     full_width = FALSE,
#     font_size = 8
#   ) |>
#   column_spec(1, width = "22%") |>  # Topography
#   column_spec(2, width = "28%") |>  # Alternative_name
#   column_spec(3, width = "28%") |>  # MESH_URI
#   column_spec(4, width = "30%") |>  # Definition_NALT
#   # column_spec(5, width = "30%") |>  # Definition_MESH
#   # column_spec(6, width = "22%") |>  # Alternative_name
#   row_spec(0, bold = TRUE, font_size = 12) |>  # Header row styling
#   scroll_box(width = "740px", height = "1000px")

  eye_table %>% bind_rows() |> 
  gt::gt() %>%
  gt::tab_options(table.font.size = 12) %>%
  gt::cols_width(Topography ~ gt::pct(22),
                 Alternative_name ~ gt::pct(20),
                 Definition_NALT ~ gt::pct(30)) %>%
  gt::tab_style(
    style = cell_text(size = px(15), weight = "bold"),
    locations = cells_column_labels())

```

##### Nose

```{r necropsy nose, echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE, cache=F}

# source("Necropsy_anatomy/Necropsy_topology_nose.R")

# saveRDS(nose_table, "Necropsy_anatomy/Tables for manual/nose.RDS")
nose_table <- readRDS("Necropsy_anatomy/Tables for manual/nose.RDS")

# nose_table |> 
#   mutate(Topography = stringr::str_to_sentence(Topography)) |> 
#   mutate(Alternative_name = stringr::str_to_sentence(Alternative_name)) |> 
#   mutate(Definition_NALT = ifelse(Definition_NALT == "", NA, Definition_NALT))|> 
#   arrange(Topography) |> 
#   kable("html") |>
#   kable_styling(
#     bootstrap_options = c("striped", "hover"),
#     full_width = FALSE,
#     font_size = 8) |>
#   column_spec(1, width = "22%") |>  # Topography
#   column_spec(2, width = "28%") |>  # Alternative_name
#   column_spec(3, width = "28%") |>  # MESH_URI
#   column_spec(4, width = "30%") |>  # Definition_NALT
#   # column_spec(5, width = "30%") |>  # Definition_MESH
#   # column_spec(6, width = "22%") |>  # Alternative_name
#   row_spec(0, bold = TRUE, font_size = 12) |>  # Header row styling
#   scroll_box(width = "740px", height = "1000px")

  nose_table %>% bind_rows() |> 
  gt::gt() %>%
  gt::tab_options(table.font.size = 12) %>%
  gt::cols_width(Topography ~ gt::pct(22),
                 Alternative_name ~ gt::pct(20),
                 Definition_NALT ~ gt::pct(30)) %>%
  gt::tab_style(
    style = cell_text(size = px(15), weight = "bold"),
    locations = cells_column_labels())

```

##### Ears

```{r necropsy ears, echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE, cache=F}

# source("Necropsy_anatomy/Necropsy_topology_ear.R")

# saveRDS(ear_table, "Necropsy_anatomy/Tables for manual/ear.RDS")
ear_table <- readRDS("Necropsy_anatomy/Tables for manual/ear.RDS")


# ear_table |> 
#   mutate(Topography = stringr::str_to_sentence(Topography)) |> 
#   mutate(Alternative_name = stringr::str_to_sentence(Alternative_name)) |> 
#   mutate(Definition_NALT = ifelse(Definition_NALT == "", NA, Definition_NALT)) |> 
#   arrange(Topography) |> 
#   kable("html") |>
#   kable_styling(
#     bootstrap_options = c("striped", "hover"),
#     full_width = FALSE,
#     font_size = 8) |>
#   column_spec(1, width = "22%") |>  # Topography
#   column_spec(2, width = "28%") |>  # Alternative_name
#   column_spec(3, width = "28%") |>  # MESH_URI
#   column_spec(4, width = "30%") |>  # Definition_NALT
#   # column_spec(5, width = "30%") |>  # Definition_MESH
#   # column_spec(6, width = "22%") |>  # Alternative_name
#   row_spec(0, bold = TRUE, font_size = 12) |>  # Header row styling
#   scroll_box(width = "740px", height = "1000px")

  ear_table %>% bind_rows() |> 
  gt::gt() %>%
  gt::tab_options(table.font.size = 12) %>%
  gt::cols_width(Topography ~ gt::pct(22),
                 Alternative_name ~ gt::pct(20),
                 Definition_NALT ~ gt::pct(30)) %>%
  gt::tab_style(
    style = cell_text(size = px(15), weight = "bold"),
    locations = cells_column_labels())

```


##### Sinuses

```{r necropsy sinuses, echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE, cache=F}

# source("Necropsy_anatomy/Necropsy_topology_sinuses.R")

# saveRDS(sinuses_table, "Necropsy_anatomy/Tables for manual/sinuses.RDS")
sinuses_table <- readRDS("Necropsy_anatomy/Tables for manual/sinuses.RDS")


# sinuses_table |> 
#   mutate(Topography = stringr::str_to_sentence(Topography)) |> 
#   mutate(Alternative_name = stringr::str_to_sentence(Alternative_name)) |> 
#   mutate(Definition_NALT = ifelse(Definition_NALT == "", NA, Definition_NALT)) |> 
#   arrange(Topography) |> 
#   kable("html") |>
#   kable_styling(
#     bootstrap_options = c("striped", "hover"),
#     full_width = FALSE,
#     font_size = 8) |>
#   column_spec(1, width = "22%") |>  # Topography
#   column_spec(2, width = "28%") |>  # Alternative_name
#   column_spec(3, width = "28%") |>  # MESH_URI
#   column_spec(4, width = "30%") |>  # Definition_NALT
#   # column_spec(5, width = "30%") |>  # Definition_MESH
#   # column_spec(6, width = "22%") |>  # Alternative_name
#   row_spec(0, bold = TRUE, font_size = 12) |>  # Header row styling
#   scroll_box(width = "740px", height = "1000px")

  sinuses_table %>% bind_rows() |> 
  gt::gt() %>%
  gt::tab_options(table.font.size = 12) %>%
  gt::cols_width(Topography ~ gt::pct(22),
                 Alternative_name ~ gt::pct(20),
                 Definition_NALT ~ gt::pct(30)) %>%
  gt::tab_style(
    style = cell_text(size = px(15), weight = "bold"),
    locations = cells_column_labels())

```


##### Integument

```{r necropsy integument, echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE, cache=F}

# source("Necropsy_anatomy/Necropsy_topology_integument.R")

# saveRDS(integument_table, "Necropsy_anatomy/Tables for manual/integument.RDS")
integument_table <- readRDS("Necropsy_anatomy/Tables for manual/integument.RDS")

integument_table<-
rbind(
integument_table,
c("Eyelashes", NA, NA, NA)) |> 
  mutate(Topography = stringr::str_to_sentence(Topography)) |> 
  mutate(Alternative_name = stringr::str_to_sentence(Alternative_name)) |> 
  mutate(Definition_NALT = ifelse(Definition_NALT == "", NA, Definition_NALT)) |> 
  arrange(Topography) #|> 
  # kable("html") |>
  # kable_styling(
  #   bootstrap_options = c("striped", "hover"),
  #   full_width = FALSE,
  #   font_size = 8) |>
  # column_spec(1, width = "22%") |>  # Topography
  # column_spec(2, width = "28%") |>  # Alternative_name
  # column_spec(3, width = "28%") |>  # MESH_URI
  # column_spec(4, width = "30%") |>  # Definition_NALT
  # # column_spec(5, width = "30%") |>  # Definition_MESH
  # # column_spec(6, width = "22%") |>  # Alternative_name
  # row_spec(0, bold = TRUE, font_size = 12) |>  # Header row styling
  # scroll_box(width = "740px", height = "1000px")

integument_table %>% bind_rows() |> 
  gt::gt() %>%
  gt::tab_options(table.font.size = 12) %>%
  gt::cols_width(Topography ~ gt::pct(22),
                 Alternative_name ~ gt::pct(20),
                 Definition_NALT ~ gt::pct(30)) %>%
  gt::tab_style(
    style = cell_text(size = px(15), weight = "bold"),
    locations = cells_column_labels())

```


##### Skin

```{r necropsy skin, echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE, cache=F}

# source("Necropsy_anatomy/Necropsy_topology_skin.R")

# saveRDS(skin_table, "Necropsy_anatomy/Tables for manual/skin.RDS")
skin_table <- readRDS("Necropsy_anatomy/Tables for manual/skin.RDS")

skin_table<-
rbind(
  skin_table,
  c("Scar", NA, NA, NA))|> 
  mutate(Topography = stringr::str_to_sentence(Topography)) |> 
  mutate(Alternative_name = stringr::str_to_sentence(Alternative_name)) |> 
  mutate(Definition_NALT = ifelse(Definition_NALT == "", NA, Definition_NALT)) |> 
  arrange(Topography) #|> 
  # kable("html") |>
  # kable_styling(
  #   bootstrap_options = c("striped", "hover"),
  #   full_width = FALSE,
  #   font_size = 8) |>
  # column_spec(1, width = "22%") |>  # Topography
  # column_spec(2, width = "28%") |>  # Alternative_name
  # column_spec(3, width = "28%") |>  # MESH_URI
  # column_spec(4, width = "30%") |>  # Definition_NALT
  # column_spec(5, width = "30%") |>  # Definition_MESH
  # column_spec(6, width = "22%") |>  # Alternative_name
  # row_spec(0, bold = TRUE, font_size = 12) |>  # Header row styling
  # scroll_box(width = "740px", height = "1000px")

skin_table %>% bind_rows() |> 
  gt::gt() %>%
  gt::tab_options(table.font.size = 12) %>%
  gt::cols_width(Topography ~ gt::pct(22),
                 Alternative_name ~ gt::pct(20),
                 Definition_NALT ~ gt::pct(30)) %>%
  gt::tab_style(
    style = cell_text(size = px(15), weight = "bold"),
    locations = cells_column_labels())
```


##### Subcutaneous Fat

```{r necropsy subcutaneous fat, echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE, cache=F}

# source("Necropsy_anatomy/Necropsy_topology_subcutaneous _fat.R")

# saveRDS(sc_fat_table, "Necropsy_anatomy/Tables for manual/fat.RDS")
sc_fat_table <- readRDS("Necropsy_anatomy/Tables for manual/fat.RDS")

# 
# sc_fat_table |> 
#   mutate(Topography = stringr::str_to_sentence(Topography)) |> 
#   mutate(Alternative_name = stringr::str_to_sentence(Alternative_name)) |> 
#   mutate(Definition_NALT = ifelse(Definition_NALT == "", NA, Definition_NALT)) |> 
#   arrange(Topography) |> 
#   kable("html") |>
#   kable_styling(
#     bootstrap_options = c("striped", "hover"),
#     full_width = FALSE,
#     font_size = 8) |>
#   column_spec(1, width = "22%") |>  # Topography
#   column_spec(2, width = "28%") |>  # Alternative_name
#   column_spec(3, width = "28%") |>  # MESH_URI
#   column_spec(4, width = "30%") |>  # Definition_NALT
#   # column_spec(5, width = "30%") |>  # Definition_MESH
#   # column_spec(6, width = "22%") |>  # Alternative_name
#   row_spec(0, bold = TRUE, font_size = 12) |>  # Header row styling
#   scroll_box(width = "740px", height = "1000px")

sc_fat_table %>% bind_rows() |> 
  gt::gt() %>%
  gt::tab_options(table.font.size = 12) %>%
  gt::cols_width(Topography ~ gt::pct(22),
                 Alternative_name ~ gt::pct(20),
                 Definition_NALT ~ gt::pct(30)) %>%
  gt::tab_style(
    style = cell_text(size = px(15), weight = "bold"),
    locations = cells_column_labels())

```

##### Muskuloskeletal System


```{r necropsy muskuloskeletal, echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE, cache=F}

# source("Necropsy_anatomy/Necropsy_topology_muscleskeletal _fat.R")

# saveRDS(muscle_table, "Necropsy_anatomy/Tables for manual/muscle.RDS")
muscle_table <- readRDS("Necropsy_anatomy/Tables for manual/muscle.RDS")

# 
# muscle_table |> 
#   mutate(Topography = stringr::str_to_sentence(Topography)) |> 
#   mutate(Alternative_name = stringr::str_to_sentence(Alternative_name)) |> 
#   mutate(Definition_NALT = ifelse(Definition_NALT == "", NA, Definition_NALT)) |> 
#   arrange(Topography) |> 
#   kable("html") |>
#   kable_styling(
#     bootstrap_options = c("striped", "hover"),
#     full_width = FALSE,
#     font_size = 8) |>
#   column_spec(1, width = "22%") |>  # Topography
#   column_spec(2, width = "28%") |>  # Alternative_name
#   column_spec(3, width = "28%") |>  # MESH_URI
#   column_spec(4, width = "30%") |>  # Definition_NALT
#   # column_spec(5, width = "30%") |>  # Definition_MESH
#   # column_spec(6, width = "22%") |>  # Alternative_name
#   row_spec(0, bold = TRUE, font_size = 12) |>  # Header row styling
#   scroll_box(width = "740px", height = "1000px")

muscle_table %>% bind_rows() |> 
  gt::gt() %>%
  gt::tab_options(table.font.size = 12) %>%
  gt::cols_width(Topography ~ gt::pct(22),
                 Alternative_name ~ gt::pct(20),
                 Definition_NALT ~ gt::pct(30)) %>%
  gt::tab_style(
    style = cell_text(size = px(15), weight = "bold"),
    locations = cells_column_labels())

```

##### Body Cavities 

```{r necropsy body cavities, echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE, cache=F}

# source("Necropsy_anatomy/Necropsy_topology_body_cavities.R")

# saveRDS(body_cavities_table, "Necropsy_anatomy/Tables for manual/body.RDS")
body_cavities_table <- readRDS("Necropsy_anatomy/Tables for manual/body.RDS")

# body_cavities_table |> 
#   mutate(Topography = stringr::str_to_sentence(Topography)) |> 
#   mutate(Alternative_name = stringr::str_to_sentence(Alternative_name)) |> 
#   mutate(Definition_NALT = ifelse(Definition_NALT == "", NA, Definition_NALT)) |> 
#   arrange(Topography) |> 
#   kable("html") |>
#   kable_styling(
#     bootstrap_options = c("striped", "hover"),
#     full_width = FALSE,
#     font_size = 8) |>
#   column_spec(1, width = "22%") |>  # Topography
#   column_spec(2, width = "28%") |>  # Alternative_name
#   column_spec(3, width = "28%") |>  # MESH_URI
#   column_spec(4, width = "30%") |>  # Definition_NALT
#   # column_spec(5, width = "30%") |>  # Definition_MESH
#   # column_spec(6, width = "22%") |>  # Alternative_name
#   row_spec(0, bold = TRUE, font_size = 12) |>  # Header row styling
#   scroll_box(width = "740px", height = "1000px")

body_cavities_table %>% bind_rows() |> 
  gt::gt() %>%
  gt::tab_options(table.font.size = 12) %>%
  gt::cols_width(Topography ~ gt::pct(22),
                 Alternative_name ~ gt::pct(20),
                 Definition_NALT ~ gt::pct(30)) %>%
  gt::tab_style(
    style = cell_text(size = px(15), weight = "bold"),
    locations = cells_column_labels())

```

##### Cardiovascular System

```{r necropsy cardio, echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE, cache=F}

# source("Necropsy_anatomy/Necropsy_topology_cardiovascular.R")

# saveRDS(circulatory_table, "Necropsy_anatomy/Tables for manual/heart.RDS")
circulatory_table <- readRDS("Necropsy_anatomy/Tables for manual/heart.RDS")


# circulatory_table |> 
#   mutate(Topography = stringr::str_to_sentence(Topography)) |> 
#   mutate(Alternative_name = stringr::str_to_sentence(Alternative_name)) |> 
#   mutate(Definition_NALT = ifelse(Definition_NALT == "", NA, Definition_NALT)) |> 
#   arrange(Topography) |> 
#   kable("html") |>
#   kable_styling(
#     bootstrap_options = c("striped", "hover"),
#     full_width = FALSE,
#     font_size = 8) |>
#   column_spec(1, width = "22%") |>  # Topography
#   column_spec(2, width = "28%") |>  # Alternative_name
#   column_spec(3, width = "28%") |>  # MESH_URI
#   column_spec(4, width = "30%") |>  # Definition_NALT
#   # column_spec(5, width = "30%") |>  # Definition_MESH
#   # column_spec(6, width = "22%") |>  # Alternative_name
#   row_spec(0, bold = TRUE, font_size = 12) |>  # Header row styling
#   scroll_box(width = "740px", height = "1000px")


circulatory_table %>% bind_rows() |> 
  gt::gt() %>%
  gt::tab_options(table.font.size = 12) %>%
  gt::cols_width(Topography ~ gt::pct(22),
                 Alternative_name ~ gt::pct(20),
                 Definition_NALT ~ gt::pct(30)) %>%
  gt::tab_style(
    style = cell_text(size = px(15), weight = "bold"),
    locations = cells_column_labels())


```

##### Respiratory System

```{r necropsy respiratory, echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE, cache=F}

# source("Necropsy_anatomy/Necropsy_topology_respiratory.R")

# saveRDS(resp_table, "Necropsy_anatomy/Tables for manual/respiratory.RDS")
resp_table <- readRDS("Necropsy_anatomy/Tables for manual/respiratory.RDS")

# resp_table |> 
#   mutate(Topography = stringr::str_to_sentence(Topography)) |> 
#   mutate(Alternative_name = stringr::str_to_sentence(Alternative_name)) |> 
#   mutate(Definition_NALT = ifelse(Definition_NALT == "", NA, Definition_NALT)) |> 
#   arrange(Topography) |> 
#   kable("html") |>
#   kable_styling(
#     bootstrap_options = c("striped", "hover"),
#     full_width = FALSE,
#     font_size = 8) |>
#   column_spec(1, width = "22%") |>  # Topography
#   column_spec(2, width = "28%") |>  # Alternative_name
#   column_spec(3, width = "28%") |>  # MESH_URI
#   column_spec(4, width = "30%") |>  # Definition_NALT
#   # column_spec(5, width = "30%") |>  # Definition_MESH
#   # column_spec(6, width = "22%") |>  # Alternative_name
#   row_spec(0, bold = TRUE, font_size = 12) |>  # Header row styling
#   scroll_box(width = "740px", height = "1000px")

resp_table %>% bind_rows() |> 
  gt::gt() %>%
  gt::tab_options(table.font.size = 12) %>%
  gt::cols_width(Topography ~ gt::pct(22),
                 Alternative_name ~ gt::pct(20),
                 Definition_NALT ~ gt::pct(30)) %>%
  gt::tab_style(
    style = cell_text(size = px(15), weight = "bold"),
    locations = cells_column_labels())


```


##### Gastrointestinal System

```{r necropsy gastrointestinal, echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE, cache=F}

# source("Necropsy_anatomy/Necropsy_topology_digestive.R")

# saveRDS(digestive_table, "Necropsy_anatomy/Tables for manual/gastro.RDS")
digestive_table <- readRDS("Necropsy_anatomy/Tables for manual/gastro.RDS")

# digestive_table |> 
#   mutate(Topography = stringr::str_to_sentence(Topography)) |> 
#   mutate(Alternative_name = stringr::str_to_sentence(Alternative_name)) |> 
#   mutate(Definition_NALT = ifelse(Definition_NALT == "", NA, Definition_NALT)) |> 
#   arrange(Topography) |> 
#   kable("html") |>
#   kable_styling(
#     bootstrap_options = c("striped", "hover"),
#     full_width = FALSE,
#     font_size = 8) |>
#   column_spec(1, width = "22%") |>  # Topography
#   column_spec(2, width = "28%") |>  # Alternative_name
#   column_spec(3, width = "28%") |>  # MESH_URI
#   column_spec(4, width = "30%") |>  # Definition_NALT
  # column_spec(5, width = "30%") |>  # Definition_MESH
  # column_spec(6, width = "22%") |>  # Alternative_name
  # row_spec(0, bold = TRUE, font_size = 12) |>  # Header row styling
  # scroll_box(width = "740px", height = "1000px")

digestive_table %>% bind_rows() |> 
  gt::gt() %>%
  gt::tab_options(table.font.size = 12) %>%
  gt::cols_width(Topography ~ gt::pct(22),
                 Alternative_name ~ gt::pct(20),
                 Definition_NALT ~ gt::pct(30)) %>%
  gt::tab_style(
    style = cell_text(size = px(15), weight = "bold"),
    locations = cells_column_labels())


```

##### Urinary Tract

```{r necropsy urinary, echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE, cache=F}

# source("Necropsy_anatomy/Necropsy_topology_urinary.R")

# saveRDS(urinary_table, "Necropsy_anatomy/Tables for manual/urinary.RDS")
urinary_table <- readRDS("Necropsy_anatomy/Tables for manual/urinary.RDS")

# urinary_table |> 
#   mutate(Topography = stringr::str_to_sentence(Topography)) |> 
#   mutate(Alternative_name = stringr::str_to_sentence(Alternative_name)) |> 
#   mutate(Definition_NALT = ifelse(Definition_NALT == "", NA, Definition_NALT)) |> 
#   arrange(Topography) |> 
#   kable("html") |>
#   kable_styling(
#     bootstrap_options = c("striped", "hover"),
#     full_width = FALSE,
#     font_size = 8) |>
#   column_spec(1, width = "22%") |>  # Topography
#   column_spec(2, width = "28%") |>  # Alternative_name
#   column_spec(3, width = "28%") |>  # MESH_URI
#   column_spec(4, width = "30%") |>  # Definition_NALT
#   # column_spec(5, width = "30%") |>  # Definition_MESH
#   # column_spec(6, width = "22%") |>  # Alternative_name
#   row_spec(0, bold = TRUE, font_size = 12) |>  # Header row styling
#   scroll_box(width = "740px", height = "1000px")

urinary_table %>% bind_rows() |> 
  gt::gt() %>%
  gt::tab_options(table.font.size = 12) %>%
  gt::cols_width(Topography ~ gt::pct(22),
                 Alternative_name ~ gt::pct(20),
                 Definition_NALT ~ gt::pct(30)) %>%
  gt::tab_style(
    style = cell_text(size = px(15), weight = "bold"),
    locations = cells_column_labels())
```

##### Reproductive System

```{r necropsy reproductvie, echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE, cache=F}

# source("Necropsy_anatomy/Necropsy_topology_repro.R")

# saveRDS(repro_table, "Necropsy_anatomy/Tables for manual/repro.RDS")
repro_table <- readRDS("Necropsy_anatomy/Tables for manual/repro.RDS")


# repro_table |> 
#   mutate(Topography = stringr::str_to_sentence(Topography)) |> 
#   mutate(Alternative_name = stringr::str_to_sentence(Alternative_name)) |> 
#   mutate(Definition_NALT = ifelse(Definition_NALT == "", NA, Definition_NALT)) |> 
#   arrange(Topography) |> 
#   kable("html") |>
#   kable_styling(
#     bootstrap_options = c("striped", "hover"),
#     full_width = FALSE,
#     font_size = 8) |>
#   column_spec(1, width = "22%") |>  # Topography
#   column_spec(2, width = "28%") |>  # Alternative_name
#   column_spec(3, width = "28%") |>  # MESH_URI
#   column_spec(4, width = "30%") |>  # Definition_NALT
#   # column_spec(5, width = "30%") |>  # Definition_MESH
#   # column_spec(6, width = "22%") |>  # Alternative_name
#   row_spec(0, bold = TRUE, font_size = 12) |>  # Header row styling
#   scroll_box(width = "740px", height = "1000px")

repro_table %>% bind_rows() |> 
  gt::gt() %>%
  gt::tab_options(table.font.size = 12) %>%
  gt::cols_width(Topography ~ gt::pct(22),
                 Alternative_name ~ gt::pct(20),
                 Definition_NALT ~ gt::pct(30)) %>%
  gt::tab_style(
    style = cell_text(size = px(15), weight = "bold"),
    locations = cells_column_labels())
```


##### Lymphatic System

```{r necropsy lymphatic, echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE, cache=F}

# source("Necropsy_anatomy/Necropsy_topology_lymph.R")
# 
# saveRDS(lympha_table, "Necropsy_anatomy/Tables for manual/lymph.RDS")
lympha_table <- readRDS("Necropsy_anatomy/Tables for manual/lymph.RDS")


# lympha_table |> 
#   mutate(Topography = stringr::str_to_sentence(Topography)) |> 
#   mutate(Alternative_name = stringr::str_to_sentence(Alternative_name)) |> 
#   mutate(Definition_NALT = ifelse(Definition_NALT == "", NA, Definition_NALT)) |> 
#   arrange(Topography) |> 
#   kable("html") |>
#   kable_styling(
#     bootstrap_options = c("striped", "hover"),
#     full_width = FALSE,
#     font_size = 8) |>
#   column_spec(1, width = "22%") |>  # Topography
#   column_spec(2, width = "28%") |>  # Alternative_name
#   column_spec(3, width = "28%") |>  # MESH_URI
#   column_spec(4, width = "30%") |>  # Definition_NALT
#   # column_spec(5, width = "30%") |>  # Definition_MESH
#   # column_spec(6, width = "22%") |>  # Alternative_name
#   row_spec(0, bold = TRUE, font_size = 12) |>  # Header row styling
#   scroll_box(width = "740px", height = "1000px")

lympha_table %>% bind_rows() |> 
  gt::gt() %>%
  gt::tab_options(table.font.size = 12) %>%
  gt::cols_width(Topography ~ gt::pct(22),
                 Alternative_name ~ gt::pct(20),
                 Definition_NALT ~ gt::pct(30)) %>%
  gt::tab_style(
    style = cell_text(size = px(15), weight = "bold"),
    locations = cells_column_labels())


```

##### Endocrine System

```{r necropsy endocrine, echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE, cache=F}

# source("Necropsy_anatomy/Necropsy_topology_endocrine.R")

# saveRDS(endo_table, "Necropsy_anatomy/Tables for manual/endo.RDS")
endo_table <- readRDS("Necropsy_anatomy/Tables for manual/endo.RDS")


# endo_table |> 
#   mutate(Topography = stringr::str_to_sentence(Topography)) |> 
#   mutate(Alternative_name = stringr::str_to_sentence(Alternative_name)) |> 
#   mutate(Definition_NALT = ifelse(Definition_NALT == "", NA, Definition_NALT)) |> 
#   arrange(Topography) |> 
#   kable("html") |>
#   kable_styling(
#     bootstrap_options = c("striped", "hover"),
#     full_width = FALSE,
#     font_size = 8) |>
#   column_spec(1, width = "22%") |>  # Topography
#   column_spec(2, width = "28%") |>  # Alternative_name
#   column_spec(3, width = "28%") |>  # MESH_URI
#   column_spec(4, width = "30%") |>  # Definition_NALT
#   # column_spec(5, width = "30%") |>  # Definition_MESH
#   # column_spec(6, width = "22%") |>  # Alternative_name
#   row_spec(0, bold = TRUE, font_size = 12) |>  # Header row styling
#   scroll_box(width = "740px", height = "1000px")

endo_table %>% bind_rows() |> 
  gt::gt() %>%
  gt::tab_options(table.font.size = 12) %>%
  gt::cols_width(Topography ~ gt::pct(22),
                 Alternative_name ~ gt::pct(20),
                 Definition_NALT ~ gt::pct(30)) %>%
  gt::tab_style(
    style = cell_text(size = px(15), weight = "bold"),
    locations = cells_column_labels())

```


##### Nervous System

```{r necropsy nervous, echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE, cache=F}

# source("Necropsy_anatomy/Necropsy_topology_nervous.R")
# 
# saveRDS(nervo_table, "Necropsy_anatomy/Tables for manual/nervous.RDS")
nervo_table <- readRDS("Necropsy_anatomy/Tables for manual/nervous.RDS")


# nervo_table |> 
#   mutate(Topography = stringr::str_to_sentence(Topography)) |> 
#   mutate(Alternative_name = stringr::str_to_sentence(Alternative_name)) |> 
#   mutate(Definition_NALT = ifelse(Definition_NALT == "", NA, Definition_NALT)) |> 
#   arrange(Topography) |> 
#   kable("html") |>
#   kable_styling(
#     bootstrap_options = c("striped", "hover"),
#     full_width = FALSE,
#     font_size = 8) |>
#   column_spec(1, width = "22%") |>  # Topography
#   column_spec(2, width = "28%") |>  # Alternative_name
#   column_spec(3, width = "28%") |>  # MESH_URI
#   column_spec(4, width = "30%") |>  # Definition_NALT
#   # column_spec(5, width = "30%") |>  # Definition_MESH
#   # column_spec(6, width = "22%") |>  # Alternative_name
#   row_spec(0, bold = TRUE, font_size = 12) |>  # Header row styling
#   scroll_box(width = "740px", height = "1000px")

nervo_table %>% bind_rows() |> 
  gt::gt() %>%
  gt::tab_options(table.font.size = 12) %>%
  gt::cols_width(Topography ~ gt::pct(22),
                 Alternative_name ~ gt::pct(20),
                 Definition_NALT ~ gt::pct(30)) %>%
  gt::tab_style(
    style = cell_text(size = px(15), weight = "bold"),
    locations = cells_column_labels())

```

### Environmental Source

**An Environmental Source is a unit in space where Specimens that cannot be associated with a Group Source, an Animal Source, or Arthropod Source can be collected from** (e.g., the site where feces of an unknown animal are found). 

Key attributes of a **Environmental Source** include:

* Environmental Source ID
* Environmental Source Cross Reference ID
* Environment Type (e.g., pond, grassland, landfill)

#### Environmental Source Record

An **Environmental Source Record** is the site where environmental **Specimens** are collected from at time *t*. Many types of abiotic or biotic tissue can be obtained from the same **Environmental Source Record**, for example, a specific site in a pond at time *t* where water, sediment, and air are obtained. Moreover, **Environmental Source Record** can contain **Diagnostics**, for example, the turbidity of the water at the site. 

Key attributes of a **Environmental Source Record** include:

* Environmental Source Record ID
* Record Number

**If multiple tissue types are obtained a time t from an Environmental Source and each of them is associated with a particular active collection effort, then there will be  multiple Collection Points under the same Event containing the same Environmental Source. In this case it is key to properly enter the Environmental Source Record IDs**.

### Arthropod Source

**An Arthropod Source is a unit in space where arthropods exist and they can be taken from** (e.g., a household in the forest where traps can be set). 

Key attributes of a **Arthropod Source** include:

* Source ID
* Cross Reference ID
* Cross Reference ID Origin

**If the interest of the Surveillance Activity is at the arthropod individual level (e.g., butterflies with problems in their wings or with parasites), then the user should consider these arthropods as Animal Sources. Arthropods from Animals Sources (attached ticks, lice, fleas, mites) are Specimens (see next section) from those Animal Source and not Arthropod Sources**.


#### Arthropod Source Record

**An Arthropod Source Record represents the set of arthropods collected from an Arthropod Source at time *t* using a specific method**. For example, the set of mosquitoes  at time *t* (**Arthropod Source Record**) from a site in the forest (**Arthropod Source**) captured using a CO₂ trap (**Collection Point**). This means that mosquitoes obtained using CO₂ traps and BG traps (**Collection Points**) at an **Arthropod Source** at time *t* belong to the same **Arthropod Source Record**.


Key attributes of a **Arthropod Source Record** include:

* Arthropod Source Record Number
* Arthropod Species
* Number by age, sex, and condition (females only)

**Similar to Environmental Sources, a key distinction between Arthropod Sources and Group or Animal Sources is that an Event can contain multiple Collection Points (e.g., each trap) containing the same Arthropod Record Record**.

**A key characteristic of an Arthropods Source Record is that "species" is a property of the Source for Group and Animal Sources, while for Arthropod Sources, there can be several species under the Source Record level**. For example, mosquitoes collected from a specific site (**Source**) using CO₂ traps (**Collection Point**) will be identified and counted by species in the **Arthropod Source Record**. 

If an arthropod **Collection** attempt at time *t* fails, the **Collection** itself is recorded, and an empty **Arthropod Source Record** is added.
